Exploration

### 2018-03-12 exploration.Rmd
added code for fitting IPM based on sampled cells

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 500
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 50  # ncells in IPM matrix
tmax <- 30
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=100, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage & IPM matrix objects
p <- sp_i$p
setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
d <- sdd.j <- vector("list", length=n.cell)
B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
nSd <- D <- p_est.i <- N_sim <- matrix(0, nrow=n.cell, ncol=tmax)
lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd[,,1:2,] <- sdd.pr

@@

#---- Realization: generate simulated data
## local growth
for(i in 1:n.cell) {
  d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
  E <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E, p, lo, hi)
  nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
  sdd[,,3,i] <- nSd[i,1] * p$p_emig * sdd[,,1,i]
}
N_sim[,1] <- n_i
## dispersal & density dependence
for(i in 1:n.cell) {
  D[i,1] <- sum(sdd[,,3,][sdd.j[[i]]])
  p_est.i[i,1] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,1]+D[i,1]),p$p_est), p$p_est)
  d[[i]] <- sim_recruits(d[[i]], p_est.i[i,1], nSd[i,1], B[i,1], D[i,1], p)
  B[i,2] <- sim_seedbank(nSd[i,1], B[i,1], D[i,1], p)
}
## iterate forward
for(k in 2:tmax) {
  for(i in 1:n.cell) {
    size.k <- d[[i]]$sizeNext[!is.na(d[[i]]$sizeNext) & d[[i]]$yr == k-1]
    if(length(size.k) > 0) {
      k.index <- (1:length(size.k)) + nrow(d[[i]])
      d[[i]][k.index, "size"] <- size.k
      d[[i]][k.index, "yr"] <- k
      E <- sim_expected(d[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
      d[[i]] <- sim_realized(d[[i]], k.index, E, p, lo, hi)
      nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
      N_sim[i,k] <- length(size.k)
    }
    sdd[,,3,i] <- nSd[i,k] * p$p_emig * sdd[,,1,i]
  }
  for(i in 1:n.cell) {
    D[i,k] <- sum(sdd[,,3,][sdd.j[[i]]])
    p_est.i[i,k] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,k]+D[i,k]),p$p_est), p$p_est)
    d[[i]] <- sim_recruits(d[[i]], p_est.i[i,k], nSd[i,k], B[i,k], D[i,k], p)
    B[i,k+1] <- sim_seedbank(nSd[i,k], B[i,k], D[i,k], p)
  }
}

@@

sample.cells <- sample((1:n.cell)[lam.df$N.surv>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
  mx <- cbind(1, as.matrix(xx))
  xxx <- seq(0, hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,]
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(mx[,2], antilogit(mx[,1:n_z$s]%*%p$s_z + c(X$s[i,]%*%p$s_x)), col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(mx[,2], mx[,1:n_z$g]%*%p$g_z + c(X$g[i,]%*%p$g_x), col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(mx[,2], antilogit(mx[,1:n_z$fl]%*%p$fl_z + c(X$fl[i,]%*%p$fl_x)),
            col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(mx[,2], exp(mx[,1:n_z$seed]%*%p$seed_z + c(X$seed[i,]%*%p$seed_x)),
            col=2)
    plot(y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, y)
  for(i in sample.cells) {
   z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,] 
   main.i <- paste("Cell", i)
   z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
   image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
  }
}
image.plot(x, x, t(apply(Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, t(apply(Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, sens(t(apply(IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
```

@@

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- c(Mx=500, CA=20, IPM=20) # number of cells in sample
```

### Correlative: MaxEnt
```{r O_Mx}
library(dismo)
Mx.P <- sample((1:nrow(lam.df))[lam.df$lambda > 1], O_n$Mx, replace=T)
pres <- 1:nrow(lam.df) %in% Mx.P
Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
                           (MxPred==0 & lambda < 1))
ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
```
### Mechanistic: CA
```{r O_CA}

```
### Mechanistic: IPM
```{r O_IPM}

```

@@

```
### Mechanistic: IPM
```{r P_IPM}
for(i in 1:ncell) {
  # ## IPMpack: estimate slopes from d[[i]]
  # ### set up seedbank transitions
  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
  d[[i]]$number <- 1
  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
                         fl=NA, seed=NA,
                         stage=c("seedbank", "seedbank", "continuous"),
                         stageNext=c("seedbank", "continuous", "seedbank"),
                         number=c(p$s_SB * (1 - p$rcr_SB), 
                                  p$rcr_SB * p$p_est, 
                                  (1 - p$rcr_dir) * p$s_SB))
  d[[i]] <- rbind(d[[i]], seedbank)
  d[[i]]$stage <- as.factor(d[[i]]$stage)
  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
  ### construct kernels
  so <- makeSurvObj(d[[i]], surv ~ size)
  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
  fo <- makeFecObj(d[[i]],
                   Formula=c(fl~size+size2, seed~size+size2),
                   Family=c("binomial", "poisson"),
                   Transform=c("none", "none"),
                   meanOffspringSize=mean(
                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
                                       !is.na(d[[i]]$sizeNext)]),
                   sdOffspringSize=sd(
                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
                                       !is.na(d[[i]]$sizeNext)]),
                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
                                                continuous=p$rcr_dir)
                   )
  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                          correction="constant")
  D.mx2 <- makeDiscreteTrans(d[[i]], 
                             continuousToDiscreteExplanatoryVariables=1)
  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
                            minSize=lo, maxSize=hi, correction="constant")
  IPM2 <- P.mx2 + F.mx2
  lambdas2[i] <- Re(eigen(IPM2)$values[1])

  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.2, col="red")
      # lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.2, col="red")
      # lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.2, col="red")
      # lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.2, col="red")
      # lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    # w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
    yl <- range(stable.dist1)#range(c(stable.dist1, stable.dist2))
    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    # lines(y, stable.dist2, col="blue")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=F)
      xxx <- seq(0, hi*.5, length.out=100)
      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
      # lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
  }
}

```

>>>>

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 50
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 50  # ncells in IPM matrix
tmax <- 10
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=100, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage & IPM matrix objects
p <- sp_i$p
setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
IPMs <- Ps <- Fs <- Fb <- array(0, dim=c(n+1, n+1, n.cell))
Nt <- array(dim=c(n+1, n+1, n.cell, tmax))
d <- E <- sdd.j <- vector("list", length=n.cell)
B <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
nSd <- D <- p_est.i <- N_sim <- matrix(0, nrow=n.cell, ncol=tmax)
lam.t <- p_est.t <- matrix(nrow=n.cell, ncol=tmax-1)
sdd <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd[,,1:2,] <- sdd.pr

@@

#---- Realization: generate simulated data
## local growth
for(i in 1:n.cell) {
  d[[i]] <- data.frame(yr=1, size=runif(n=n_i[i], min=z.rng[1], max=z.rng[2]))
  E[[i]] <- sim_expected(d[[i]], p, map(X, ~(.[i,])), n_z)
  d[[i]] <- sim_realized(d[[i]], 1:nrow(d[[i]]), E[[i]], p, lo, hi)
  nSd[i,1] <- sum(d[[i]]$seed, na.rm=T)
  sdd[,,3,i] <- nSd[i,1] * p$p_emig * sdd[,,1,i]
}
N_sim[,1] <- n_i
## dispersal & density dependence
for(i in 1:n.cell) {
  D[i,1] <- sum(sdd[,,3,][sdd.j[[i]]])
  p_est.i[i,1] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,1]+D[i,1]),p$p_est), p$p_est)
  d[[i]] <- sim_recruits(d[[i]], p_est.i[i,1], nSd[i,1], B[i,1], D[i,1], p)
  B[i,2] <- sim_seedbank(nSd[i,1], B[i,1], D[i,1], p)
}
## iterate forward
for(k in 2:tmax) {
  for(i in 1:n.cell) {
    size.k <- d[[i]]$sizeNext[!is.na(d[[i]]$sizeNext) & d[[i]]$yr == k-1]
    if(length(size.k) > 0) {
      k.index <- (1:length(size.k)) + nrow(d[[i]])
      d[[i]][k.index, "size"] <- size.k
      d[[i]][k.index, "yr"] <- k
      E <- sim_expected(d[[i]][k.index,], p, map(X, ~(.[i,])), n_z)
      d[[i]] <- sim_realized(d[[i]], k.index, E, p, lo, hi)
      nSd[i,k] <- sum(d[[i]]$seed[k.index], na.rm=T)
      N_sim[i,k] <- length(size.k)
    }
    sdd[,,3,i] <- nSd[i,k] * p$p_emig * sdd[,,1,i]
  }
  for(i in 1:n.cell) {
    D[i,k] <- sum(sdd[,,3,][sdd.j[[i]]])
    p_est.i[i,k] <- ifelse(p$NDD, min(p$NDD_n/(nSd[i,k]+D[i,k]),p$p_est), p$p_est)
    d[[i]] <- sim_recruits(d[[i]], p_est.i[i,k], nSd[i,k], B[i,k], D[i,k], p)
    B[i,k+1] <- sim_seedbank(nSd[i,k], B[i,k], D[i,k], p)
  }
}

@@

sample.cells <- sample((1:n.cell)[lam.df$N.surv>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
  mx <- cbind(1, as.matrix(xx))
  xxx <- seq(0, hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,]
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(mx[,2], antilogit(mx[,1:n_z$s]%*%p$s_z + c(X$s[i,]%*%p$s_x)), col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), ylim=c(lo,hi),
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(mx[,2], mx[,1:n_z$g]%*%p$g_z + c(X$g[i,]%*%p$g_x), col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(mx[,2], antilogit(mx[,1:n_z$fl]%*%p$fl_z + c(X$fl[i,]%*%p$fl_x)),
            col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         ylim=range(c(exp(mx[,1:n_z$seed]%*%p$seed_z + c(X$seed[i,]%*%p$seed_x)),
                      z.t$seed), na.rm=TRUE),
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(mx[,2], exp(mx[,1:n_z$seed]%*%p$seed_z + c(X$seed[i,]%*%p$seed_x)),
            col=2)
    plot(y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(d[[i]]$sizeNext[is.na(d[[i]]$size)], col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, y)
  for(i in sample.cells) {
   z.t <- d[[i]][!is.na(d[[i]]$size) & d[[i]]$yr==tmax,] 
   main.i <- paste("Cell", i)
   z.t$y <- y[findInterval(z.t$size, y, all.inside=T)]
   totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
   image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
  }
}
image.plot(x, x, t(apply(Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, t(apply(Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, sens(t(apply(IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
```

@@

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
P.i <- which(lam.df$N.surv > 0) # presences: surviving individuals > seedlings
```

### Correlative: MaxEnt
```{r O_Mx}
library(dismo)
Mx.P <- sample(P.i, O_n$Mx, replace=T)
pres <- 1:nrow(lam.df) %in% Mx.P
Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
                           (MxPred==0 & lambda < 1))
ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
```
### Mechanistic: CA
```{r O_CA}

```
### Mechanistic: IPM
```{r O_IPM}
IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
IPM.d <- vector("list", O_n$IPM)
for(i in IPM.P) {
  IPM.d[[i]] <- d[[i]] %>% filter(yr %in% O_yr$IPM) %>%
    add_column(id.inbd=i) %>%
    full_join(env.df[i,], by="id.inbd")
}
IPM.d <- do.call(rbind, IPM.d)
p.ipm <- p
```

@@

```
### Mechanistic: IPM
```{r P_IPM}
# run chunk O_IPM first to generate observed dataset `IPM.d`

########
## fit models
########
s.m <- glm(surv ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn + 
               pOth + pDec, data=IPM.d, family="binomial")
g.m <- lm(sizeNext ~ size + I(size^2) + temp + temp2 + prec + prec2 +
              pOpn + pOth, data=IPM.d)
fl.m <- glm(fl ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn +
                pOth, data=IPM.d, family="binomial")
seed.m <- glm(seed ~ size + I(size^2) + temp + temp2 + prec + prec2 + pOpn, 
              data=IPM.d, family="poisson")
p.ipm$s_z <- coef(s.m)[1:length(p.ipm$s_z)]
p.ipm$s_x <- coef(s.m)[(1:length(p.ipm$s_x)) + length(p.ipm$s_z)]
p.ipm$g_z <- coef(g.m)[1:length(p.ipm$g_z)]
p.ipm$g_x <- coef(g.m)[(1:length(p.ipm$g_x)) + length(p.ipm$g_z)]
p.ipm$g_sig <- summary(g.m)$sigma
p.ipm$fl_z <- coef(fl.m)[1:length(p.ipm$fl_z)]
p.ipm$fl_x <- coef(fl.m)[(1:length(p.ipm$fl_x)) + length(p.ipm$fl_z)]
p.ipm$seed_z <- coef(seed.m)[1:length(p.ipm$seed_z)]
p.ipm$seed_x <- coef(seed.m)[(1:length(p.ipm$seed_x)) + length(p.ipm$seed_z)]
p.ipm$rcr_z <- c(mean(IPM.d[is.na(IPM.d$size), "sizeNext"]),
                 sd(IPM.d[is.na(IPM.d$size), "sizeNext"]))

setup_IPM_matrix(n, z.rng, buffer=0.5, discrete=1)
IPMs.f <- Ps.f <- Fs.f <- Fb.f <- array(0, dim=c(n+1, n+1, n.cell))
Nt.f <- array(dim=c(n+1, n+1, n.cell, tmax))
d.f <- E.f <- sdd.j <- vector("list", length=n.cell)
B.f <- cbind(rpois(n.cell, n_i), matrix(0, nrow=n.cell, ncol=tmax))
nSd.f <- D.f <- p_est.i.f <- N_sim.f <- matrix(0, nrow=n.cell, ncol=tmax)
lam.t.f <- p_est.t.f <- matrix(nrow=n.cell, ncol=tmax-1)
sdd.f <- array(0, dim=dim(sdd.pr) + c(0,0,1,0)) # add layer for nEmig
sdd.f[,,1:2,] <- sdd.pr

# set up environmental covariates
n_z.f <- list(s=length(p.ipm$s_z), g=length(p.ipm$g_z),
            fl=length(p.ipm$fl_z), seed=length(p.ipm$seed_z))
n_x.f <- list(s=length(p.ipm$s_x), g=length(p.ipm$g_x),
            fl=length(p.ipm$fl_x), seed=length(p.ipm$seed_x))
X.f <- list(s=as.matrix(env.df[,1:n_x.f$s]), g=as.matrix(env.df[,1:n_x.f$g]),
          fl=as.matrix(env.df[,1:n_x.f$fl]), seed=as.matrix(env.df[,1:n_x.f$seed]))

#---- Truth: use assigned slopes to fill IPM matrix
## local growth
for(i in 1:n.cell) {
  Ps.f[,,i] <- fill_P(h, y, z.i, n, p.ipm, n_z.f, n_x.f, X.f$s[i,], X.f$g[i,])
  Fb.f[,,i] <- fill_F(h, y, z.i, n, p.ipm, n_z.f, n_x.f, X.f$fl[i,], X.f$seed[i,])
  Fs.f[z.i,z.i,i] <- (1-p.ipm$p_emig) * Fb.f[z.i,z.i,i]
  Fs.f[1,z.i,i] <- (1-p.ipm$p_emig) * Fb.f[1,z.i,i]
  Fs.f[z.i,1,i] <- Fb.f[z.i,1,i]
}
## dispersal & density dependence
for(i in 1:n.cell) {
  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes
  sdd.j[[i]] <- which(sdd.f[,,2,]==sdd.i, arr.ind=TRUE) # j cells with target i
  p.ij <- sdd.f[,,1,][sdd.j[[i]]] # p(j to i | immigrant seed from j)
  Fs.f[z.i,z.i,i] <- Fs.f[z.i,z.i,i] + 
    Reduce(`+`, map2(sdd.j[[i]][,3], p.ipm$p_emig*p.ij, ~(Fb.f[z.i,z.i,.x] * .y)))
  Fs.f[1,z.i,i] <- Fs.f[1,z.i,i] + Fb.f[1,z.i,sdd.j[[i]][,3]] %*% (p.ipm$p_emig*p.ij)
  if(p.ipm$NDD) {
    Nt.f[,,i,1] <- rpois((n+1)*(n+1), n_i[i])
    F.i.f <- F.t.f <- Fs.f[,,i]/p.ipm$p_est  
    for(k in 1:(tmax-1)) {
      p_est.t.f[i,k] <- min(p.ipm$NDD_n/(sum(F.i.f[z.i,]*Nt.f[z.i,,i,k])), p.ipm$p_est)
      F.t.f[z.i,] <- F.i.f[z.i,] * p_est.t[i,k]
      IPM.t.f <- Ps.f[,,i] + F.t.f
      Nt.f[,,i,k+1] <- IPM.t.f %*% Nt.f[,,i,k]
      lam.t.f[i,k] <- Re(eigen(IPM.t.f)$values[1])
    }
  }
}
IPMs.f <- Ps.f + Fs.f
